\chapter{Referencias}
\label{Referencias}

\bibitem{bannister2001}
Bannister, M., 2001. Challenges for composites into the next millennium – a reinforcement perspective. Compos. A 31, 901–910.

\bibitem{bardy2012}
Bardy, J., Legrand, X., Crosky, A., 2012. Configuration of a genetic algorithm used to optimise fibre steering in composite laminates. Compos. Struct. 94, 2048–2056.

\bibitem{beasley1993}
Beasley, D., Bull, D.R., Martin, R.R., 1993. An overview of genetic algorithms: Part 1, fundamentals. Univ. Comput. 15 (2), 58–69.

\bibitem{crosky2006}
Crosky, A., Kelly, D., Li, R., Legrand, X., Huong, N., Ujjin, R., 2006. Improvement of bearing strength of laminated composites. Compos. Struct. 76, 260–271.

\bibitem{crosky2012}
Crosky, A., Grant, C., Kelly, D., 2012. Fiber placements. In: Nocolais, N., Borzacchiello, A., Lee, S.M. (Eds.), Wiley Encyclopedia of Composites, second ed. Wiley, Hoboken, NJ, pp. 945–950.

\bibitem{gliesche2003}
Gliesche, K., Hubner, T., Orawetz, H., 2003. Application of the tailored fibre placement (TFP) process for a local reinforcement on an “open-hole” tension plate from carbon/epoxy laminates. Compos. Sci. Technol. 63 (1), 81–88.

\bibitem{goldberg1989}
Goldberg, D.E., 1989. Genetic Algorithms in Search, Optimisation, and Machine Learning. Addison Wesley, Reading, MA.

\bibitem{grant2006}
Grant, C., 2006. Automated processes for composite aircraft structure. Ind. Rob. 33 (2), 117–121.

\bibitem{grant2010}
Grant, C., 2010. Composites automation: there’s a lot of room for growth. In: SME Aerospace \& Defense Manufacturing Yearbook 2010. SME, Dearborn.

\bibitem{grant2012}
Grant, C., 2012. Composites automation: trending smaller and robotic. Compos. World.

\bibitem{honda2011}
Honda, S., Narita, Y., 2011. Vibration design of laminated fibrous composite plates with local anisotropy induced by short fibers and curvilinear fibers. Compos. Struct. 93, 902–910.

\bibitem{kelly2001a}
Kelly, D.W., Willgoss, R., Li, R., Crosky, A., 2001a. Improvement of bearing strength of mechanically fastened composite joints using fibre steering. In: Repecka, L., Saremi, F.F. (Eds.), Proceedings of the 48th International SAMPE Symposium, Long Beach, pp. 2495–2506.

\bibitem{kelly2001b}
Kelly, D.W., Hsu, P., Asadullah, M., 2001b. Load paths and load flow in finite element analysis. Eng. Comput. 18 (1/2), 304–313.

\bibitem{legrand2006}
Legrand, X., Crosky, A., Kelly, D., Crepin, D., 2006. Optimisation of fibre steering in composite laminates using a genetic algorithm. Compos. Struct. 75 (1–4), 524–531.

\bibitem{li2002a}
Li, R., Kelly, D., Crosky, A., 2002a. Strength improvement by fibre steering around a pin loaded hole. Compos. Struct. 57, 377–383.

\bibitem{li2002b}
Li, R., Kelly, D.W., Arima, S., Willgoss, R.A., Crosky, A.G., 2002b. Fiber steering around a cut-out in a shear loaded panel. In: Rasmussen, B.M., Pilato, L.A., Kliger, H.S. (Eds.), Proceedings of the 49th International SAMPE Symposium, Long Beach, pp. 1853–1861.

\bibitem{li2006}
Li, R., Kelly, D., Crosky, A., Schoen, H., Smolich, L., 2006. Joint efficiency of fibre steered composite joint using load path trajectories. J. Compos. Mater. 40 (18), 1645–1658.

\bibitem{peters2011}
Peters, S.T., 2011. Composites Filament Winding. ASM International, Materials Park, OH.

\bibitem{sloan2009}
Sloan, J., 2009. AFP/ATL design to manufacture: bridging the gap. High Perform. Compos.

\bibitem{starr2000}
Starr, T.F., 2000. Pultrusion for Engineers. Woodhead, Cambridge.

\bibitem{tabakov2010}
Tabakov, P.Y., Walker, A., 2010. A technique for stiffness improvement by optimization of fiber steering in composite plates. Appl. Compos. Mater. 17

\bibitem{Miravete}
Miravete, Antonio. 2011. Materiales Compuestos. EDITORIAL REVERTÉ S.A.

\bibitem{Lubin}
Lubin, George. 1982. HANDBOOK OF COMPOSITES. VAN NOSTRAND REINHOLD COMPANY

\bibitem{MIL-HDBK-17-3F}
MIL-HDBK-17-3F. 1997. COMPOSITE MATERIALS HANDBOOK. DISTRIBUTION STATEMENT A.

\bibitem{AMFLSC}
Additive Manufacturing for Lightweight Structural Components. 2023. The Automated Fiber Placement Process: Design Cycle, Benefits, and Applications. 

\bibitem{SAMPE}
Society for the Advancement of Material and Process Engineering. 2011. Automated Tape Laying and Fiber Placement Technologies.

\bibitem{SAMPE}
Mongoose HybridTM - Automatic Fiber Placement Machine by Camozzi Machine Tools | DirectIndustry. (s. f.). https://www.directindustry.com/prod/camozzi-machine-tools/product-26616-2242753.html

\bibitem{SAMPE}
Coriolis Composites: the reference in automated fiber placement. (2022, 13 diciembre). Coriolis. https://www.coriolis-composites.com/

\bibitem{SAMPE}
AddComposites | Plug and Place AFP. (s. f.). Addcomposites. https://www.addcomposites.com/
\textbf{}